\dppTitle

### 2014-05-31 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on exchangeability section of dpp ms.

\subsubsection{Ordering of taxon-specific summary statistics}
As alluded to in the model description, \msb does not maintain the order of the
taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
Rather, the summary statistics are re-ordered by descending values of average
pairwise differences between the descendant populations
($\pi_b$) \cite{NeiLi1979,Huang2011}.
\change{
Thus, the summary statistics being compared when calculating the Euclidean
distance between the observed data and each simulation are often from sequence
alignments of \emph{different} taxon pairs and loci.
In other words, the compared statistics are often calulated from datasets
simulated with \emph{different} numbers and length of sequences, and
potentially different numbers of loci, and different fixed \hky model
parameters (\hkyModel{}{}) and mutation-rate (\mutationRateScalarConstant{}{})
and ploidy (\ploidyScalar{}{}) scaling parameters.}

This has the advantage of reducing the sample space of the number of
divergence models \divTimeIndexVector, but there are at least two disadvantages.
First, additional information in the data is lost.
By discarding the identity of the \npairs{} pairs of populations, all
pair-specific information about the amount of data (e.g., the number gene
copies collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) is lost.
Second, the results are more difficult to interpret, because they can be no
longer be directly associated to the taxa under study.

@@

Thus, the theoretical basis for this reshuffling is questionable.
It may be justifiable as an additional coarsening of the data,
by removing the information associated with the identity of the
population pairs.
However, I do not provide a formal proof that this re-ordering of the pairs
does not introduce bias.
One can imagine situations in which the sampling intensity (i.e., the
number of gene copies, loci, and locus length) is highly skewed across
pairs and/or there is large heterogeneity in mutational processes
(e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
In such cases, it seems possible that discarding this information could
cause bias.
Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
is to minimize Euclidean distances between simulated datasets in which the true
model has a single divergence event \cite{Huang2011}, and (2) the reordering
increases the model's tendency to infer a single divergence event
\cite{Huang2011}, it seems possible that this approach could be biasing the
method \cite{Oaks2012}.

For comparability with \msb, I maintain the re-ordering of taxon-specific
summary statistics by $\pi_b$.  However, I also allow the order to be
maintained, and ordered divergence models to be estimated.

>>>>

\subsubsection{Ordering of taxon-specific summary statistics}
As alluded to in the model description, \msb does not maintain the order of the
taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
Rather, the summary statistics are re-ordered by descending values of average
pairwise differences between the descendant populations
($\pi_b$) \cite{NeiLi1979,Huang2011}.
\change{
As a result, the summary statistics being compared when calculating the Euclidean
distance between the observed data and each simulation are often from sequence
alignments of \emph{different} taxon pairs and/or loci.
More specifically, the summary statistics calculated from the observed sequence
alignments are being compared to summary statistics calulated from datasets
simulated with potentially \emph{different} 
(1) numbers of sequences,
(2) length of sequences,
(3) numbers of loci,
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
and
(6) ploidy multipliers (\ploidyScalar{}{}).}

This has the advantage of reducing the sample space of the number of
divergence models \divTimeIndexVector, but there are at least two disadvantages.
First, additional information in the data is lost.
By discarding the identity of the \npairs{} pairs of populations, all
pair-specific information about the amount of data (e.g., the number of gene
copies collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) is lost.
Second, the results are more difficult to interpret, because they can no
longer be directly associated to the taxa under study.

@@

\change{
Thus, whenever the six parameters listed above vary across any of the taxa or
loci (i.e., nearly all empirical applications), the taxon-specific summary
statistics \alignmentSS{}{} are not exchangeable.
Hence, the reshuffling of the summary statistic vectors in not mathematically
valid.
The magnitude of the affect of this violaton of exchangeability is currently
unknown.
Huang et al.\ \cite{Huang2011} demonstrated that the reordering of the summary
statistic vectors can greatly increase the method's tendency to infer a single
divergence event.
By definition, if the summary statistic vectors were exhangeable, the
reordering would not change the likelihood or posterior (barring sampling
error).
Thus, the results of Huang et al.\ \cite{Huang2011} suggest the
reordering of the statistics could be introducing a bias.
% Thus, the theoretical basis for this reshuffling is questionable.
% It may be justifiable as an additional coarsening of the data,
% by removing the information associated with the identity of the
% population pairs.
% However, I do not provide a formal proof that this re-ordering of the pairs
% does not introduce bias.
% One can imagine situations in which the sampling intensity (i.e., the
% number of gene copies, loci, and locus length) is highly skewed across
% pairs and/or there is large heterogeneity in mutational processes
% (e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
% In such cases, it seems possible that discarding this information could
% cause bias.
% Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
% is to minimize Euclidean distances between simulated datasets in which the true
% model has a single divergence event \cite{Huang2011}, and (2) the reordering
% increases the model's tendency to infer a single divergence event
% \cite{Huang2011}, it seems possible that this approach could be biasing the
% method \cite{Oaks2012}.
}

For comparability with \msb, I maintain the re-ordering of taxon-specific
summary statistics by $\pi_b$.  However, I also allow the order to be
maintained, and ordered divergence models to be estimated.
\change{
Also, in all of the simulation-based analyses described below, the
summary statistic vectors \emph{are} exchangeable, because all of the
simulated datasets have the same 
(1) numbers of sequences,
(2) length of sequences,
(3) numbers of loci,
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
and
(6) ploidy multipliers (\ploidyScalar{}{}).}
